---
title: "cbswrangle"
author: "Job Spijker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cbswrangle}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The cbswrangle package contains a few helper functions to make
working with the cbsodataR package a little easier.


## cbsodataR

This package is an addition on the *cbsodataR* package. The
*cbsodataR* package allows for retrieving data from 
[Statistics Netherlands (CBS)](http://www.cbs.nl), which produces all official statistics of the Netherlands. 

More information about cbsodataR can be found in the accompanying
vignette of the package. use the following command to retrieve the
vignette:

```{r, eval=FALSE}
vignette("cbsodataR")
```

## Table of contents

The available tables at CBS are stored in a table of contents, the
_TOC_

The TOC must be downloaded before any other table can be downloaded.
Downloaded data is stored in the directory 'data', to separate data
and code. The datafile package takes care of this.

```{r}

library(here)
library(cbsodataR)
library(cbswrangle)
library(dplyr)

# install with remotes::install_github("jspijker/datafile")
library(datafile)
datafileInit(basedir=tempdir())

cbstoc <- getcbstoc(fname=datafile("cbstoc.rds"))

```

# downloading a table

Each table at CBS has an identifier. These identifiers can be found in
the TOC. 

```{r}
cbstoc %>% 
  select(Identifier, ShortTitle) 

```

With `getcbs` one can download a table.  
This function downloads the data and the meta data from a CBS table and
stores it to disk. It checks if the table exists using the TOC and then
checks if the file already exists on disk. If not, the data is downloaded. If
the file is already present, then the data is read from the file.

After downloading, the fieldnames of the data are renamed and the
unit of the field is added to the fieldname. A translation table is
created with the original fieldname and the fieldname created by
this function.

The code below downloads the agriculture statistics (Dutch:
landbouwtellingen) and stores the result in
_./data/cbs\_landbouwtelling.rds_


```{r}

landbouwtelling <- getcbs("80781ned",datafile("cbs_landbouwtelling.rds"))
print(names(landbouwtelling))
```


The `getcbs` function returns A list with several objects: data contains
the data.frame with data, meta contains the meta data, tocentry contains
the meta data from the TOC, keyName is the translation table, and
fieldnames is a table with filednames and grouping id's.

CBS fieldnames folow a certain hierarchilcal grouping order. For
example in the data about agriculture you find the groups farms and
animals. In each group there is the variable pigs\_amount
(varkens\_aantallen). In the farm group this is refering to the
amount of farms with pigs, in the animal group this is refering to
the amount of pigs. From the fieldname it's not clear to which
group the field belongs, use the function `getFieldParents. to
get this information.


```{r}

fieldn <- "Leghennen_aantal.1"
getFieldParents(fieldn,landbouwtelling)

```

